Modelling last glacial cycle ice dynamics in the Alps

### 2017-09-25 articles/cycle/alps-cycle.tex
Rephrase conclusions.

    This study consists in the application of the numerical ice sheet model
    PISM to ice dynamics of the last glacial cycle in the Alps, using a model
    set-up based on that previously developed and validated for the Cordilleran
    ice sheet (Sect.~\ref{sec:model}).

    Using three different palaeo-temperature forcing records (GRIP, EPICA, and
    MD01-2444, Sect.~\ref{sec:paltemp}), scaled to model the Rhine Glacier
    piedmont lobe in agreement with the mapped LGM ice margin, and two
    different palaeo-precipitation parametrizations (with and without
    precipitation reductions, Sect.~\ref{sec:palprec}), we find that only the
    EPICA palaeo-temperature record yields model results in agreement with
    geological findings, in the sense that:

@@

    \begin{itemize}
      \item Ice cover is generally underestimated in the Western Alps and
            overestimated in the Eastern Alps, indicating that east-west
            gradients in temperature or precipitation change, absent from our
            model forcing, probably controlled the LGM extent of ice cover in
            the Alps (Sect.~\ref{sec:extent}).
      \item The assymetric nature of LGM ice extent north and south of the Alps
            can be explained by the transient nature of the LGM extent without
            involving north-south gradients in temperature and precipitation
            change (Sect.~\ref{sec:extent}).
      \item The LGM ice flow pattern in the Alps is largely controlled by
            subglacial topography, but several transfluences and two
            self-sustained ice domes occured (Sect.~\ref{sec:flow}).
      \item The modelled LGM is a transient stage in which glaciers are not in
            equilibrium with climate. Timing potentially varies across the
            range (Sect.~\ref{sec:timing}).
      \item Ice thickness during the LGM is modelled to be much higher than in
            reconstructions. In average modeleld surface elevation is 861\,m
            above the Rhone Glacier trimlines, which may instead indicate an
            englacial thermal boundary (Sect.~\ref{sec:thickness}).
      \item Aline glaciers were potentially very dyamic. They responded quickly
            to climate fluctuation and some of them potentially advanced many
            times over the foreland during the last glacial cycle
            (Sect.~\ref{sec:glaciations})
    \end{itemize}

    These results are nevertheless limited by the simplicity of the climate
    forcing and surface mass balance model applied, uncertain ice physical
    model parameters. Through more detailed sensitivity studies, targeted to
    specific aspects of the last glacial cycle ice dynamics in the Alps, future
    modelling studies will certainly be able to quantify uncertainties
    associated with some of the above statements. However, we also hope that
    these conclusion will also serve as a basis for future studies of glacial
    geology in the Alps, and call for a more systematic aggregation and
    homogeneisation of glacial geological data to form a basis for model
    validation across the entire Alpine range.

>>>>

    This study consists in the application of the numerical ice sheet model
    PISM to ice dynamics of the last glacial cycle in the Alps, using a model
    set-up based on that previously developed and validated for the Cordilleran
    ice sheet (Sect.~\ref{sec:model}).
    Using three different palaeo-temperature forcing records (GRIP, EPICA, and
    MD01-2444, Sect.~\ref{sec:paltemp}), scaled to model the Rhine Glacier
    piedmont lobe in agreement with the mapped LGM ice margin, and two
    different palaeo-precipitation parametrizations (with and without
    precipitation reductions, Sect.~\ref{sec:palprec}), we find that only the
    EPICA palaeo-temperature record yields model results in agreement with
    geological findings, in the sense that:

@@

    \begin{itemize}
      \item Ice cover is generally underestimated in the Western Alps and
            overestimated in the Eastern Alps, indicating that east-west
            gradients in temperature or precipitation change, absent from our
            model forcing, probably controlled the LGM extent of ice cover in
            the Alps. The asymmetric extent north and south of the Alps
            can be explained by the transient nature of the LGM extent without
            involving north-south gradients in temperature and precipitation
            change (Sect.~\ref{sec:extent}).
      \item The LGM ice flow pattern in the Alps is largely controlled by
            subglacial topography, but several transfluences and two
            self-sustained ice domes occured (Sect.~\ref{sec:flow}).
      \item The LGM was a transient stage in which glaciers were not in
            equilibrium with climate. Timing potentially varied across the
            range due to inherent glacier dynamics (Sect.~\ref{sec:timing}).
      \item Ice thickness during the LGM is modelled to be much higher than in
            reconstructions. In average, modelled surface elevation is 861\,m
            above the Rhone Glacier trimlines, which may instead indicate an
            englacial thermal boundary (Sect.~\ref{sec:thickness}).
      \item Alpine glaciers were potentially very dynamic. They quickly
            responded to climate fluctuation and some potentially advanced many
            times over the foreland during the last glacial cycle
            (Sect.~\ref{sec:glaciations}).
    \end{itemize}

    These results are nevertheless limited by the simplicity of the climate
    forcing and surface mass balance model applied, and uncertain ice physical
    model parameters. Through more detailed sensitivity studies, targeted to
    specific aspects of the last glacial cycle ice dynamics in the Alps, future
    modelling studies will certainly be able to quantify uncertainties
    associated with some of the above statements. Nevertheless, we hope that
    these conclusions will also serve as a basis for future studies of glacial
    geology in the Alps, and call for a more systematic aggregation and
    homogenization of glacial geological data to form a basis for model
    validation across the entire Alpine range.
